Omix

### 2019-04-09 omix/omix.Rmd
updated omix Rmd and html

@@

Edited `.Renviron` to avoid "Error: memory exhausted (limit reached?)" when melting the GE dataset, as per the suggestion from [StackOverflow](https://stackoverflow.com/questions/51295402/r-on-macos-error-vector-memory-exhausted-limit-reached):å
```
cd ~
touch .Renviron
open .Renviron
# Added following line to .Renviron:
R_MAX_VSIZE=100Gb
```

```{r}
library(crunch)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(kableExtra)
library(naniar)
library(openxlsx)
library(reshape2)
library(tidyverse)
library(VennDiagram)

@@

@@

# Remove two rows that contain cell lines with two different histologies
# The histologies that are kept are verified through matching in Cellosaurus and Cell Model Passports
# KMH2, haematopoietic_and_lymphoid: https://web.expasy.org/cellosaurus/CVCL_1330, https://cellmodelpassports.sanger.ac.uk/passports/SIDM01018
# MS1, lung: https://web.expasy.org/cellosaurus/CVCL_1429, https://cellmodelpassports.sanger.ac.uk/passports/SIDM00245
ccl_info <- subset(ccl_info, CCLE_Name != "KMH2_THYROID" & CCLE_Name != "MS1_SKIN")
```

Adding TCGA labels:
```{r}
tcga_hist_converter <- read.xlsx("./data/TCGA/Histology_DepMap_Xena_04052019.xlsx", sheet = 3, startRow = 1, colNames = TRUE, rowNames = FALSE)
tcga_hist_converter$data_source <- NULL

tcga_hist_converter_depmap <- filter(tcga_hist_converter, grepl("^ACH", sample_ID))
tcga_hist_converter_tcga <- filter(tcga_hist_converter, grepl("^TCGA", sample_ID))
colnames(tcga_hist_converter_depmap) <- c("Broad_ID", "TCGA_classification")
colnames(tcga_hist_converter_tcga) <- c("sample", "TCGA_classification")

ccl_info <- merge(ccl_info, tcga_hist_converter_depmap, by = "Broad_ID", all.x = TRUE)
```

@@

```{r}
ccl_histologies <- select(ccl_info, Cell_line, Histology, TCGA_classification, CCLE_Name, Broad_ID) %>% drop_na(Cell_line)

auc_na_histologies$Histology <- ifelse(auc_na_histologies$Histology == "", NA, as.character(auc_na_histologies$Histology))
auc_na_histologies$CCLE_Name <- ifelse(auc_na_histologies$CCLE_Name  == "", NA, as.character(auc_na_histologies$CCLE_Name ))
auc_na_histologies$Broad_ID <- ifelse(auc_na_histologies$Broad_ID == "", NA, as.character(auc_na_histologies$Broad_ID))

ccl_histologies <- subset(ccl_histologies, !(Broad_ID %in% auc_na_histologies$Broad_ID))
# Bind DepMap histologies and curated histologies
ccl_histologies <- bind_rows(ccl_histologies, auc_na_histologies)
```

@@

sum(auc_all$HasCopyNumberData)
sum(auc_all$HasMutationData)
sum(auc_all$HasGeneExpressionData)
sum(auc_all$WasDrugScreened)

test <- filter(auc_all, HasMutationData == 1)[, 1:12]

pgx_ccls <- filter(auc_all, WasDrugScreened == 1)
crispr_ccls <- filter(auc_all, WasCRISPRScreened == 1)
unscreened_ccls <- filter(auc_all, WasCRISPRScreened == 0 & WasDrugScreened == 0)

ge_ccls <- filter(auc_all, HasGeneExpressionData == 1)
cn_ccls <- filter(auc_all, HasCopyNumberData == 1)
mut_ccls <- filter(auc_all, HasMutationData == 1)

all_omics_ccls <- intersect(intersect(unique(ge_ccls$Cell_line), unique(cn_ccls$Cell_line)), unique(mut_ccls$Cell_line))
sum(filter(auc_all, Cell_line %in% all_omics_ccls)$WasDrugScreened)

ccle_ccls <- filter(auc_all, InCCLE == 1)
ctrp_ccls <- filter(auc_all, InCTRP == 1)
gdsc_ccls <- filter(auc_all, InGDSC == 1)

@@

```{r, eval=FALSE}
futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_pgx_crispr_overlap <- venn.diagram(list("CRISPR" = unique(crispr_ccls$Cell_line), "Drug Screen" = unique(pgx_ccls$Cell_line), "Unscreened" = unique(unscreened_ccls$Cell_line)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened via small-molecule and CRISPR", main.fontface = "bold", lty = rep("blank", 3), fill = c("mediumpurple1", "turquoise4", "slategray3"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./pozdeyev_unmatched/venn_pgx_crispr_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_pgx_crispr_overlap)
dev.off()

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_ccl_overlap <- venn.diagram(list("CCLE" = unique(ccle_ccls$Cell_line), "CTRP" = unique(ctrp_ccls$Cell_line), "GDSC" = unique(gdsc_ccls$Cell_line)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened in each dataset", main.fontface = "bold", lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./pozdeyev_unmatched/venn_ccl_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_ccl_overlap)
dev.off()

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_omics_overlap <- venn.diagram(list("Gene Expression" = unique(ge_ccls$Cell_line), "Copy Number" = unique(cn_ccls$Cell_line), "Mutation" = unique(mut_ccls$Cell_line)), resolution = 5000, main = "Overlapping omic annotations in cancer cell lines (CCLs)", main.fontface = "bold", lty = rep("blank", 3), fill = c("yellow2", "orange", "tomato"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./pozdeyev_unmatched/venn_omics_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_omics_overlap)
dev.off()

bar_ccl_per_hist <- ggplot(data = unique(select(auc_all, Cell_line, Histology))) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  scale_y_continuous(breaks = seq(0, 300, by = 25), labels = seq(0, 300, by = 25)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology across all cell lines") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_ccl_per_hist
ggsave("./pozdeyev_unmatched/barplot_ccl_per_histology.png", bar_ccl_per_hist, device = "png", dpi = 450, width = 12, height = 6, units = "in")

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_drugs_overlap <- venn.diagram(list("CCLE" = unique(ccle_poz$Drug_name), "CTRP" = unique(ctrp_poz$Drug_name), "GDSC" = unique(gdsc_poz$Drug_name)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened in each dataset", main.fontface = "bold", lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./pozdeyev_unmatched/venn_drugs_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_drugs_overlap)
dev.off()

bar_ccl_per_hist_per_db <- ggplot(data = unique(select(auc, Database, Cell_line, Histology))) +
  facet_wrap(~ Database) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  # coord_cartesian(ylim = c(0, 130)) +
  scale_y_continuous(breaks = seq(0, 160, by = 20), labels = seq(0, 160, by = 20)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology, grouped by database") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_ccl_per_hist_per_db
ggsave("./pozdeyev_unmatched/barplot_ccl_per_histology_per_db.png", bar_ccl_per_hist_per_db, device = "png", dpi = 450, width = 12, height = 8, units = "in")

bar_ccl_per_hist <- ggplot(data = unique(select(auc_all, Cell_line, Histology))) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  scale_y_continuous(breaks = seq(0, 375, by = 25), labels = seq(0, 375, by = 25)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology across all annotated cell lines") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_ccl_per_hist
ggsave("./pozdeyev_unmatched/barplot_ccl_per_histology.png", bar_ccl_per_hist, device = "png", dpi = 450, width = 12, height = 8, units = "in")

histo_aucic50_per_db <- ggplot(data = auc) +
  facet_wrap(~ Database, scales = "free_y") +
  geom_histogram(aes(x = AUC_IC50, fill = Database), color = "black", alpha = 0.7, binwidth = 0.025, boundary = 0, closed = "left") +
  scale_fill_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  guides(fill = "none") +
  labs(x = "AUC IC50", y = "Frequency", title = "Distributions of AUC IC50 values for all drugs in each dataset (unmatched)")
histo_aucic50_per_db
ggsave("./pozdeyev_unmatched/histogram_aucic50_per_db.png", histo_aucic50_per_db, device = "png", dpi = 450, width = 12, height = 6, units = "in")
```

@@

merged_auc_v20190319.1.csv.gz

merged_auc_v20190319.2.csv.gz:
- NAs in Histology column (excluding most of the cell lines indicated as being from a "collaborator" in CTRP) hand-curated by searches in DepMap, COSMIC, Cellosaurus, ATCC, canSAR Black, and Cell Model Passports

merged_auc_v20190327.3.csv.gz:
- Reformatted to have only unique cell lines in rows
- Cancer cell lines screened in more than one database have multiple AUC columns (column names are formatted as Drug_DatabaseName)
- Added information on binary presence/absencee of omic annotations for each cell line (i.e. mutations, copy number, gene expression)
- Added binary columns indicating presence in CCLE, CTRP, GDSC, drug screens in general (in 1, 2, or 3), and the Avana CRISPR screen
- Added cell lines that have at least one type of omic data but are not in the CRISPR or pharmacogenomic screens

merged_auc_v20190402.4.csv.gz:
- Fixed duplicate cell line rows derived from manually curating some of the CCL histologies that had been added in merged_auc_v20190319.2.csv.gz
- Fixed HasMutationData, HasCopyNumberData, and HasGeneExpression data columns
- Reordered columns

merged_auc_v20190408.5.csv.gz:
- Removed duplicate cell line entries for KMH2 and MS1 that had incorrect histology entries
- Added TCGA classifications for all cell lines possible (classifications were manually curated by Kaiqiao Li and Sally Claridge by combining CCLE and GDSC TCGA annotations and two-fold manual comparison)
- Changed CCLE_ccl_name column to CCLE_Name

# COMPARE TO TCGA

@@

Quality check: cluster TCGA samples and cell lines—verify that same tumor types cluster together rather than clustering by cell lines and tumor samples (Kaiqiao)

Calculate Pearson correlations between data sets and create a heatmap (also look at Barretina et al. supplement for PCA and mutation load correlations)

```{r, eval=FALSE}
# DepMap gene expression
ge_depmap <- read.delim("./../../htscreens_giant_files/omix/GE/GE_DepMap_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(ge_depmap)[colnames(ge_depmap) == "cancer"] <- "TCGA_classification"
# Mean center scale
ge_depmap[, 4:ncol(ge_depmap)] <- apply(ge_depmap[, 4:ncol(ge_depmap)], MARGIN = 2, FUN = scale, center = TRUE, scale = TRUE)
# Melt
ge_depmap_melt <- melt(ge_depmap, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(ge_depmap)[4:length(colnames(ge_depmap))], variable.name = "Gene", value.name = "TPM_scaled")
# Average scaled TPM per gene per TCGA classification
ge_depmap_avg <- ge_depmap_melt %>% group_by(TCGA_classification, Gene) %>% summarise(Mean_TPM = mean(TPM_scaled))
# Make wide
ge_depmap_wide <- dcast(data = ge_depmap_avg, formula = Gene ~ TCGA_classification, fun.aggregate = mean)
rownames(ge_depmap_wide) <- ge_depmap_wide$Gene
ge_depmap_wide$Gene <- NULL
colnames(ge_depmap_wide) <- paste(colnames(ge_depmap_wide), "DepMap", sep = "_")

@@

ge_wide <- bind_cols(ge_depmap_wide, ge_tcga_wide)

ge_corr <- ge_wide %>% as.matrix %>% cor(method = "pearson") %>% as.data.frame %>% rownames_to_column(var = "var1") %>% gather(var2, value, -var1)
ge_corr <- filter(ge_corr, !(grepl("DepMap", var1) & grepl("DepMap", var2)) & !(grepl("TCGA", var1) & grepl("TCGA", var2)))
ge_corr_wide <- dcast(ge_corr, formula = var1 ~ var2, fun.aggregate = mean)
# ge_wide_corr <- apply(ge_wide_corr, 2, function(x) gsub("_TCGA|_DepMap","", x))

# ge4corr <- merge(ge_depmap_avg, ge_tcga_avg, by = c("TCGA_classification", "Gene"), all = TRUE)
# ge_corr <- ge4corr %>% group_by(TCGA_classification) %>% summarize(corr = cor.test(y = Mean_TPM_TCGA, x = Mean_TPM_DepMap, method = "pearson", use = "complete.obs")$estimate)
```

# (OLD) Data management

>>>>

# NOTES

- Quality check: cluster TCGA samples and cell lines—verify that same tumor types cluster together rather than clustering by cell lines and tumor samples (Kaiqiao)
- Calculate Pearson correlations between DepMap and TCGA feature sets and create heatmaps (CN, GE, mutation frequency), see Barretina et al. (2016) supplement for example methods and plots
- Look at Barretina et al. supplement for PCA and mutation load correlations)

@@

Edited `.Renviron` to avoid "Error: memory exhausted (limit reached?)" when melting the GE dataset, as per the suggestion from [StackOverflow](https://stackoverflow.com/questions/51295402/r-on-macos-error-vector-memory-exhausted-limit-reached):
```
cd ~
touch .Renviron
open .Renviron
# Added following line to .Renviron:
R_MAX_VSIZE=100Gb
```

```{r}
library(circlize)
library(ComplexHeatmap)
library(crunch)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(kableExtra)
library(naniar)
library(openxlsx)
library(reshape2)
library(tidyverse)
library(VennDiagram)

@@

```{r}
unfactorize <- function(df){
  for(i in which(sapply(df, class) == "factor")) {
    df[[i]] = as.character(df[[i]])
  }
  return(df)
}
```

@@

# Remove two rows that contain cell lines with two different histologies
# The histologies that are kept are verified through matching in Cellosaurus and Cell Model Passports
# KMH2, haematopoietic_and_lymphoid: https://web.expasy.org/cellosaurus/CVCL_1330, https://cellmodelpassports.sanger.ac.uk/passports/SIDM01018
# MS1, lung: https://web.expasy.org/cellosaurus/CVCL_1429, https://cellmodelpassports.sanger.ac.uk/passports/SIDM00245
ccl_info <- subset(ccl_info, CCLE_Name != "KMH2_THYROID" & CCLE_Name != "MS1_SKIN")

# Add TCGA labels
tcga_hist_converter_depmap <- filter(read.xlsx("./data/TCGA/Histology_DepMap_Xena_04052019.xlsx", sheet = 3, startRow = 1, colNames = TRUE, rowNames = FALSE), data_source == "DepMap")
colnames(tcga_hist_converter_depmap) <- c("Broad_ID", "TCGA_classification", "data_source")

tcga_hist_converter_tcga <- filter(read.xlsx("./data/TCGA/Histology_DepMap_Xena_04052019.xlsx", sheet = 3, startRow = 1, colNames = TRUE, rowNames = FALSE), data_source == "Xena_TCGA")
colnames(tcga_hist_converter_tcga) <- c("sample", "TCGA_classification", "data_source")

ccl_info <- merge(ccl_info, tcga_hist_converter_depmap, by = "Broad_ID", all.x = TRUE)

# Using Pozdeyev filtering
ccl_histologies <- select(ccl_info, Cell_line, Histology, TCGA_classification, CCLE_Name, Broad_ID) %>% drop_na(Cell_line)

auc_na_histologies$Histology <- ifelse(auc_na_histologies$Histology == "", NA, as.character(auc_na_histologies$Histology))
auc_na_histologies$CCLE_Name <- ifelse(auc_na_histologies$CCLE_Name  == "", NA, as.character(auc_na_histologies$CCLE_Name ))
auc_na_histologies$Broad_ID <- ifelse(auc_na_histologies$Broad_ID == "", NA, as.character(auc_na_histologies$Broad_ID))
auc_na_histologies <- unfactorize(auc_na_histologies)

ccl_histologies <- subset(ccl_histologies, !(Broad_ID %in% auc_na_histologies$Broad_ID))
ccl_histologies <- unfactorize(ccl_histologies)
# Bind DepMap histologies and curated histologies
ccl_histologies <- bind_rows(ccl_histologies, auc_na_histologies)
```

@@

@@

auc_all_summary <- t(auc_all %>% select(InGDSC, InCTRP, InCCLE, HasCopyNumberData, HasMutationData, HasGeneExpressionData, WasDrugScreened, WasCRISPRScreened) %>% summarise_each(funs(count = sum)))
auc_all_summary

@@

```{r, eval=FALSE}
# Data subsetting for easier tracking throughout plots
pgx_ccls <- filter(auc_all, WasDrugScreened == 1)
crispr_ccls <- filter(auc_all, WasCRISPRScreened == 1)
unscreened_ccls <- filter(auc_all, WasCRISPRScreened == 0 & WasDrugScreened == 0)
ge_ccls <- filter(auc_all, HasGeneExpressionData == 1)
cn_ccls <- filter(auc_all, HasCopyNumberData == 1)
mut_ccls <- filter(auc_all, HasMutationData == 1)
ccle_ccls <- filter(auc_all, InCCLE == 1)
ctrp_ccls <- filter(auc_all, InCTRP == 1)
gdsc_ccls <- filter(auc_all, InGDSC == 1)

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_pgx_crispr_overlap <- venn.diagram(list("CRISPR" = unique(crispr_ccls$Cell_line), "Drug Screen" = unique(pgx_ccls$Cell_line), "Unscreened" = unique(unscreened_ccls$Cell_line)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened via small-molecule and CRISPR", main.fontface = "bold", lty = rep("blank", 3), fill = c("mediumpurple1", "turquoise4", "slategray3"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./figures/venn_pgx_crispr_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_pgx_crispr_overlap)
dev.off()

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_ccl_overlap <- venn.diagram(list("CCLE" = unique(ccle_ccls$Cell_line), "CTRP" = unique(ctrp_ccls$Cell_line), "GDSC" = unique(gdsc_ccls$Cell_line)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened in each dataset", main.fontface = "bold", lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./figures/venn_ccl_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_ccl_overlap)
dev.off()

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_omics_overlap <- venn.diagram(list("Gene Expression" = unique(ge_ccls$Cell_line), "Copy Number" = unique(cn_ccls$Cell_line), "Mutation" = unique(mut_ccls$Cell_line)), resolution = 5000, main = "Overlapping omic annotations in cancer cell lines (CCLs)", main.fontface = "bold", lty = rep("blank", 3), fill = c("yellow2", "orange", "tomato"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./figures/venn_omics_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_omics_overlap)
dev.off()

bar_ccl_per_hist <- ggplot(data = unique(select(auc_all, Cell_line, Histology))) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  scale_y_continuous(breaks = seq(0, 300, by = 25), labels = seq(0, 300, by = 25)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology across all cell lines") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_ccl_per_hist
ggsave("./figures/barplot_ccl_per_histology.png", bar_ccl_per_hist, device = "png", dpi = 450, width = 12, height = 6, units = "in")

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_drugs_overlap <- venn.diagram(list("CCLE" = unique(ccle_poz$Drug_name), "CTRP" = unique(ctrp_poz$Drug_name), "GDSC" = unique(gdsc_poz$Drug_name)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened in each dataset", main.fontface = "bold", lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./figures/venn_drugs_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_drugs_overlap)
dev.off()

bar_ccl_per_hist_per_db <- ggplot(data = unique(select(auc, Database, Cell_line, Histology))) +
  facet_wrap(~ Database) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  # coord_cartesian(ylim = c(0, 130)) +
  scale_y_continuous(breaks = seq(0, 160, by = 20), labels = seq(0, 160, by = 20)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology, grouped by database") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_ccl_per_hist_per_db
ggsave("./figures/barplot_ccl_per_histology_per_db.png", bar_ccl_per_hist_per_db, device = "png", dpi = 450, width = 12, height = 8, units = "in")

bar_ccl_per_hist <- ggplot(data = unique(select(auc_all, Cell_line, Histology))) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  scale_y_continuous(breaks = seq(0, 375, by = 25), labels = seq(0, 375, by = 25)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology across all annotated cell lines") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_ccl_per_hist
ggsave("./figures/barplot_ccl_per_histology.png", bar_ccl_per_hist, device = "png", dpi = 450, width = 12, height = 8, units = "in")

histo_aucic50_per_db <- ggplot(data = auc) +
  facet_wrap(~ Database, scales = "free_y") +
  geom_histogram(aes(x = AUC_IC50, fill = Database), color = "black", alpha = 0.7, binwidth = 0.025, boundary = 0, closed = "left") +
  scale_fill_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  guides(fill = "none") +
  labs(x = "AUC IC50", y = "Frequency", title = "Distributions of AUC IC50 values for all drugs in each dataset (unmatched)")
histo_aucic50_per_db
ggsave("./figures/histogram_aucic50_per_db.png", histo_aucic50_per_db, device = "png", dpi = 450, width = 12, height = 6, units = "in")
```

@@

- merged_auc_v20190319.1.csv.gz
- merged_auc_v20190319.2.csv.gz
    - NAs in Histology column (excluding most of the cell lines indicated as being from a "collaborator" in CTRP) hand-curated by searches in DepMap, COSMIC, Cellosaurus, ATCC, canSAR Black, and Cell Model Passports
    - merged_auc_v20190327.3.csv.gz
    - Reformatted to have only unique cell lines in rows
    - Cancer cell lines screened in more than one database have multiple AUC columns (column names are formatted as Drug_DatabaseName)
    - Added information on binary presence/absencee of omic annotations for each cell line (i.e. mutations, copy number, gene expression)
    - Added binary columns indicating presence in CCLE, CTRP, GDSC, drug screens in general (in 1, 2, or 3), and the Avana CRISPR screen
    - Added cell lines that have at least one type of omic data but are not in the CRISPR or pharmacogenomic screens
- merged_auc_v20190402.4.csv.gz
    - Fixed duplicate cell line rows derived from manually curating some of the CCL histologies that had been added in merged_auc_v20190319.2.csv.gz
    - Fixed HasMutationData, HasCopyNumberData, and HasGeneExpression data columns
    - Reordered columns
- merged_auc_v20190408.5.csv.gz
    - Removed duplicate cell line entries for KMH2 and MS1 that had incorrect histology entries
    - Added TCGA classifications for all cell lines possible (classifications were manually curated by Kaiqiao Li and Sally Claridge by combining CCLE and GDSC TCGA annotations and two-fold manual comparison)
    - Changed CCLE_ccl_name column to CCLE_Name

# DEPMAP/TCGA: Heatmaps

@@

Cancer cell lines and tumor samples are directly incomparable by nature, so the correlation matrices are asymmetric. The top left shows how well the features of the DepMap cancer cell lines correlate with those of the TCGA tumor samples, and the bottom right shows the reverse. The diagonal of the heatmap shows the comparison of matched TCGA classifications.

## GE

Data wrangling and processing:
```{r ge_heatmap_data, eval=FALSE}
# DepMap gene expression
ge_depmap <- read.delim("./../../htscreens_giant_files/omix/GE/GE_DepMap_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(ge_depmap)[colnames(ge_depmap) == "cancer"] <- "TCGA_classification"
# Mean center scale
ge_depmap[, 4:ncol(ge_depmap)] <- apply(ge_depmap[, 4:ncol(ge_depmap)], MARGIN = 2, FUN = scale, center = TRUE, scale = TRUE)
# Melt
ge_depmap_melt <- melt(ge_depmap, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(ge_depmap)[4:length(colnames(ge_depmap))], variable.name = "Gene", value.name = "TPM_scaled")
# Average scaled TPM per gene per TCGA classification
ge_depmap_avg <- ge_depmap_melt %>% group_by(TCGA_classification, Gene) %>% summarise(Mean_TPM = mean(TPM_scaled))
# Make wide
ge_depmap_wide <- dcast(data = ge_depmap_avg, formula = Gene ~ TCGA_classification, fun.aggregate = mean)
rownames(ge_depmap_wide) <- ge_depmap_wide$Gene
ge_depmap_wide$Gene <- NULL
colnames(ge_depmap_wide) <- paste(colnames(ge_depmap_wide), "DepMap", sep = "_")

@@

# Correlate all against all
ge_wide <- bind_cols(ge_depmap_wide, ge_tcga_wide)
ge_corr <- ge_wide %>% as.matrix %>% cor(method = "spearman") %>% as.data.frame %>% rownames_to_column(var = "var1") %>% gather(var2, value, -var1)
# Filter self-comparisons and eliminate duplicate correlations
ge_corr <- filter(ge_corr, !(grepl("DepMap", var1) & grepl("DepMap", var2)) & !(grepl("TCGA", var1) & grepl("TCGA", var2)) & grepl("TCGA", var2))
# Data management
colnames(ge_corr) <- c("DepMap", "TCGA", "r")
ge_corr <- data.frame(apply(ge_corr, 2, function(x) gsub("_TCGA|_DepMap","", x)))
ge_corr$r <- as.numeric(as.character(ge_corr$r))
# Set up coprrelation dataframe for plotting
ge_corr_wide <- dcast(ge_corr, DepMap ~ TCGA)
rownames(ge_corr_wide) <- ge_corr_wide$DepMap
ge_corr_wide$DepMap <- NULL
```

Pseudo-figure caption: This is comparison of gene expression profiles between DepMap cancer cell lines and TCGA tumor samples, both classified by TCGA project. In each dataset and for each TCGA classification, gene expression (in TPM) for each of the 5,000 most variable genes shared between the two datasets were averaged across samples. The dataset averages per TCGA classification were correlated against each other, and these Pearson correlations (r) between TCGA and DepMap gene expression feature sets are shown in the heatmap.

Make heatmaps:
```{r ge_heatmap, eval=FALSE, fig.height=4, fig.width=3.75}
col_fun <- colorRamp2(c(-0.5, 0, 0.5, 1), c("black", "red3", "yellow", "white"))
ge_ht4order <- Heatmap(ge_corr_wide, cluster_rows = FALSE)

ge_ht_legend_list <- list(legend_direction = "horizontal", title_position = "lefttop", legend_width = unit(1.5, "in"), border = "black")

ge_ht <- Heatmap(ge_corr_wide, 
                 name = "Expression correlation",
                 col = col_fun, 
                 rect_gp = gpar(col = "black"),
                 heatmap_legend_param = ge_ht_legend_list,
                 row_order = rev(column_order(ge_ht4order)), 
                 show_column_dend = FALSE, 
                 cluster_rows = FALSE, 
                 row_names_side = "left",
                 row_title = "DepMap cancer cell lines",
                 column_title = "TCGA tumor samples", 
                 column_title_side = "bottom")
png(filename = "./figures/ge_depmap_tcga_heatmap_spearman.png", width = 7.5, height = 8, units = "in", res = 500)
draw(ge_ht, heatmap_legend_side = "top")
dev.off()
```

## CN

Data wrangling and processing:

**DO NOT RUN UNTIL ADDED TCGA TO CN AND GE DATASETS**
```{r cn_heatmap_data, eval=FALSE}
# TCGA copy number
cn_tcga <- read.delim("./data/CN/CN_GDC_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(cn_tcga)[colnames(cn_tcga) == "cancer"] <- "TCGA_classification"
# Melt
cn_tcga_melt <- melt(cn_tcga, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(cn_tcga)[4:length(colnames(cn_tcga))], variable.name = "Gene", value.name = "CN")
# Average scaled TPM per gene per TCGA classification
cn_tcga_avg <- cn_tcga_melt %>% group_by(TCGA_classification, Gene) %>% summarise(Mean_CN = mean(CN))
# Make wide
cn_tcga_wide <- dcast(data = cn_tcga_avg, formula = Gene ~ TCGA_classification, fun.aggregate = mean)
rownames(cn_tcga_wide) <- cn_tcga_wide$Gene
cn_tcga_wide$Gene <- NULL
colnames(cn_tcga_wide) <- paste(colnames(cn_tcga_wide), "TCGA", sep = "_")

## 18Q4

# DepMap copy number (18Q4)
cn_depmap <- read.delim("./data/CN/CN_DepMap_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(cn_depmap)[colnames(cn_depmap) == "cancer"] <- "TCGA_classification"
# Melt
cn_depmap_melt <- melt(cn_depmap, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(cn_depmap)[4:length(colnames(cn_depmap))], variable.name = "Gene", value.name = "CN")
# Average CN levels per gene per TCGA classification
cn_depmap_avg <- cn_depmap_melt %>% group_by(TCGA_classification, Gene) %>% summarise(Mean_CN = mean(CN))
# Make wide
cn_depmap_wide <- dcast(data = cn_depmap_avg, formula = Gene ~ TCGA_classification, fun.aggregate = mean)
rownames(cn_depmap_wide) <- cn_depmap_wide$Gene
cn_depmap_wide$Gene <- NULL
colnames(cn_depmap_wide) <- paste(colnames(cn_depmap_wide), "DepMap", sep = "_")

# Correlate all against all (18Q4)
cn_wide <- bind_cols(cn_depmap_wide, cn_tcga_wide)
cn_corr <- cn_wide %>% as.matrix %>% cor(method = "spearman") %>% as.data.frame %>% rownames_to_column(var = "var1") %>% gather(var2, value, -var1)
# Filter self-comparisons and eliminate duplicate correlations
cn_corr <- filter(cn_corr, !(grepl("DepMap", var1) & grepl("DepMap", var2)) & !(grepl("TCGA", var1) & grepl("TCGA", var2)) & grepl("TCGA", var2))
# Data management
colnames(cn_corr) <- c("DepMap", "TCGA", "r")
cn_corr <- data.frame(apply(cn_corr, 2, function(x) gsub("_TCGA|_DepMap","", x)))
cn_corr$r <- as.numeric(as.character(cn_corr$r))
# Set up coprrelation dataframe for plotting
cn_corr_wide <- dcast(cn_corr, DepMap ~ TCGA)
rownames(cn_corr_wide) <- cn_corr_wide$DepMap
cn_corr_wide$DepMap <- NULL

## 19Q1
